*Prochlorococcus marinus* responses to light and oxygen

### 2024-05-24 Output/PLOSRound2/ResponsetoReviewers.txt

Reviewer #1: This manuscript presents a helpful expansion of our knowledge of the ecological tolerance 
ranges of several Prochlorococcus ecotypes. It is especially interesting in its implications for range 
expansion of this genus under potential future conditions, although this angle could be expanded on in 
my opinion, as I explain below.

Specific Comments:

The abstract seems to omit any discussion of the conclusions/bigger picture of the study.

Line 55: LLII/III and LLIV are not the “oldest” clades, but the most basal or early branching.

We have edited the text in the revised manuscript: "Clades LLII and LLIII, including cultured strain SS120, are grouped together as early branching phylogenetic lineage diversifying in the P. marinus radiation, with a preference for low light."
Line: XXX

@@

Line 369: The authors invoke “limitations in the annotation process” to explain the lack of evidence for 
MED4 ATP synthase. What does this mean? The MED4 genome surely contains ATP synthase genes.

We agree that the MED4 genome includes genes encoding ATP Synthase, but in the Ocean Protein Portal database no ATP Synthase peptides are annotated as coming from MED4. This likely reflects an annotation assignment issue because the ATP Synthase proteins are highly conserved, so assignments to strains or clades of Prochlorococcus may be ambiguous.
We have edited the text in the revised manuscript: "Though present in the genome of MED4, the absence of proteins annotated as ATP Synthase complex for Clade HLI, compared to annotated detections of ATP Synthase across the other three clades suggests limitations in the annotation process for highly conserved protein sequences."
Line: XXX

@@

Thank you for this suggestion. 
We have edited the text:
Line: XXX

Figure S4 — It is very hard to see the fit lines other than the pooled fit. Consider using different colors 
and solid lines since this is online.

Table 1: The trees described in the legend should be included in the supplemental material.

Thank you for bringing this to our attention. We clarified the wording on line XXXX of the revised manuscripted. "Protein homologies between *Prochlorococcus* and *Synechocystis* were extracted from @bonisteelStrainSpecificDifferences2018."

@@

L757: The authors should specify that this is a DNA repair ligase, not the DNA ligase that seals Okazaki 
fragments during DNA replication (assuming this is the case).

Thank you for your comment, the revised text now reads "DNA repair ligase".  
Line: XXX

L 758: Photolyase should not have been relevant in the lab experiments because there is no UV light 
emitted (AFAIK) by the bulbs used in the Multicultivator. Also, it is unlikely that the absence of these two 
genes is sufficient to explain why these strains can’t grow at the ocean’s surface.

Thank you. The revised manuscript now reads: "The absence of genes encoding deoxyribodipyrimidine photolyase and DNA ligase (ATP) in P. marinus MIT9313 and P. marinus SS120 may, in part, explain why these two strains cannot tolerate growth under full [O2] and high light, found at the ocean surface."
Line: XXX

@@

My primary concern with the manuscript comes from the way the results are framed throughout. The 
authors describe results as informing about the growth of Prochlorococcus across a matrix of different 
conditions. However, my interpretation of the methods is that the authors took from a maintenance 
culture kept under standard growth conditions and then immediately transferred them into the 
Multicultivator to look at responses. This is of course completely valid, but extensive experience by 
many in the field (e.g. described in Moore et al 2007 Limn. Oceanog. Methods and elsewhere) is that 
these cells typically require multiple generations to acclimate to new growth conditions. I read these 
data as more reflecting the immediate response of Prochlorococcus to environmental changes. It is 
quite conceivable that some of these strains may exhibit a different growth rate than what you report if 
allowed to gradually acclimate to their new condition rather than directly move them. I would suggest 
that the framing and interpretation of the paper should reflect this. I also have no idea whether this 
might have influenced the surprising observations of MIT9313 growth at relatively high PAR. Global 
change scenarios would presumably give cells many generations to physiologically acclimate to 
differences (as well as to evolutionarily adapt).

Thank you for highlighting this oversight. 
We added the following in the materials and methods section of the revised manuscript:
"Each tube containing 70 mL of Pro99 media was inoculated with 10 mL of growing maintenance culture approximately 16 to 22 hours
before the start of each growth experiment. The tubes containing the cultures were placed in the Multicultivator water bath set at 22°C, 
sparged with the experimental [O2], and kept in the dark for approximately 12 hours until the photoregime of a sinusoidal photoperiod 
commenced the following morning such that peak PAR occurred at noon each day."
Line: XXX

We added the following in the results and discussion section of the revised manuscript:
"While we implemented measures to minimize the shock to cultures from exposure to experimental growth conditions by inoculating them the day
 before the experiment began and employing a sinusoidal photoperiod with a gradual increase in PAR exposure, we acknowledge that cells typically
 need multiple generations to acclimate to new growth conditions (Moore et al. 2007)."
Line: XXX

@@

40: Despite _its_ small cell size (?)

We have edited the text:
Line: XXX

56: “Affinity” for low light? Consider rephrasing

We have edited the text in the revised manuscript: "Clades LLII and LLIII, including cultured strain SS120, are grouped together as early branching phylogenetic lineage diversifying in the *P. marinus* radiation, with a preference for low light."
Line: XXX

@@

84: Follett et al 2022 PNAS showed that temp is not the only factor influencing poleward issues - 
ecological interactions may further impact poleward shift during warming

Thank you for  bringing this to our attention. We have added the following text to the revised manuscript: "Follett et al 2022, however, model interactions of heterotrophic bacteria  that may influence latitudinal expansion of *P. marinus*."
Line XXX:

@@

199: Could you please state more about why you are confident in this assignment? What validation or 
metric leads you to this?

Thank you for noting that this needs further clarification. 
We have edited the paragraph for clarity in the revised manuscript adding: "While the accuracy of strain specific protein annotation is noted due to highly conserved common proteins, the two step approach of peptide-to-spectrum matching using deep paired metagenomics identify proteins at the clade classifications."
Lines: XXX

@@

209: It seems overly general to blame the culturing difficulties solely on having reduced genomes – in 
this context, it is the lack of just one gene (catalase) that matters, and reduced genomes reflect an 
outcome of the ultimate suite of selective pressures acting on these systems.

Thank you for your comment. We have edited the text in the revised manuscript: "Prochlorococcus remain challenging to culture at high densities, partially due to their dependence upon mutualistic heterotrophic bacteria to detoxify
reactive oxygen species [48,49]." 
Line XXX:

@@

363: Could the authors comment somewhere on why you took this proteomic approach to examine the 
baseline distribution of Pro across O2 and depth, as opposed to (or at least in concert with) the 
extensive metagenomic, amplicon, qPCR, etc datasets of ecotype abundances available?

Thank you for your comment. We used proteomics as we have experience with cyanobacteria protein and have work flows in place.
We have edited the text in the revised manuscript: "Our analyses utilized a proteomics dataset; however, alternate approaches, such as metagenomics or metatranscriptomics, could have been employed to analyze ecotype abundances using the TARA Oceans [@pesantOpenScienceResources2015] dataset."
Line: XXX

@@

407: growth -> grew?

We have edited the text: "The growth rate for Prochlorococcus marinus MED4, clade HLI, under 250 µM O2, increased with higher imposed PAR and longer photoperiods (Fig 3A), across all spectral wavebands."
Line XXX:

@@

428: How many replicates were carried out for most experiments? Perhaps I just missed it, but please be 
sure this information is clear in the methods

Thank you for this comment. In this work we used 1 to 3 replicates depending on the experimental conditions. Specific number of replicates (shown as small dots in Figures (3 - 5)).
We have edited the text in the revised manuscript:" Consistent absence of growth of some strains under some levels of photoperiod, PAR, or [O~2~] meant we completed 268 growth factor treatment combinations with 1 to 3 replicates."
Line: XXX

@@

431, and generally throughout the paper: Would the authors care to comment on any connection 
between the differences in blue vs red light growth and the underlying amount of energy input from 
these different wavelengths?

XXX Energy- blue have shorter WL so high energy, red longer WL so lower energy??? XXXX

Fig 4 and related figures: It might be useful to plot the specific measured points within the countour plot

Thank you for your suggestion. We have combined the growth rate figure with the GAM model for each strain. We believe this will aid the reader with interpreting both plots.

### 2024-05-24 Code/Manuscript_O2.Rmd
edits

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD
PSI Multicultivators (Fig \@ref(fig:LabeledMC); PSI, Drásov, Czech
Republic). Each multicultivator individually controls 8 tubes at a
common temperature of 22°C. Each tube containing 70 mL of Pro99 media
was inoculated with 10 mL of growing maintenance culture. In a factorial
matrix design, each tube was then subject to an individual combination
of sinusoidal photoperiod (4, 8, 12, 16 h); reaching a peak PAR (30, 90,
180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White
LED, 660 nm, 450 nm). [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed
by bubbling tubes with varying ratios of air and Nitrogen (N~2~), with
consistent 0.05% of Carbon Dioxide (CO~2~) gas, delivered through a 0.2
μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc.,
Kingston, Ontario, Canada). [O~2~] *in situ* was verified using oxygen
optodes (PyroScience, Germany) inserted into tubes for real-time
measurements, with a temperature probe in the bath of the bioreactor to
correct [O~2~] measures for temperature fluctuations. In addition, the
Pyroscience software corrected [O~2~] based on the salinity of the media
(32 ppt). The flow rate of the gas mixture was controlled, but
variations in bubbling speed, PAR and culture density affected the
[O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as
2.5 µM hereafter), was achieved by sparging with a gas mixture
containing 99.95% N~2~ and 0.05% CO~2~. An intermediate [O~2~] of 10 -
25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas
mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of
200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging
with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

Guided in part by the evidence of ocean distributions of proteins from
*Prochlorococcus* we set up a matrix of photoperiods, PAR, spectral
bands, and [O~2~] to approximate current, and potential future,
latitudinal, depth and seasonal niches for *Prochlorococcus* strains. As
mentioned, growth under red light could prove mechanistically
informative [@murphyPhotoinactivationPhotosystemII2017] to factors
limiting *Prochlorococcus* growth, we therefore included the red
spectral waveband even though it is not representative of
*Prochlorococcus* niches. Although *Prochlorococcus* is currently
limited to a narrow range of surface photoperiods, potential poleward
latitudinal expansions, in combination with attenuation of light with
depth, mean *Prochlorococcus* may potentially encounter a wide range of
photoperiods. Our growth rate determinations generally agree with those
from Moore *et al*. [@mooreComparativePhysiologySynechococcus1995], for
white LED and 250 µM O~2~, but our study is, to our knowledge, the first
to analyze the interactive growth responses of *Prochlorococcus* strains
to varying [O~2~], spectral wavebands and photoperiods.

@@

The GAM model in Fig \@ref(fig:SS120PARGAM)B summarizes growth responses
of SS120 to red or blue peak PAR and photoperiod, across the
2 imposed oxygen concentrations. Under 250 µM O~2~, Fig \@ref(fig:SS120PARGAM)B showed highest growth rates below blue light PAR of
50 µmol photons m^-2^ s^-1^ and photoperiods between 8 and 12 h,
indicated by the contour line labeled 0.18 d^-1^ (representing the
90^th^ percentile of achieved growth rate). Under 250 µM O~2~ SS120 is
constrained to deeper ocean waters through its intolerance of higher
blue PAR levels. These findings align with Moore *et al*.
[@mooreComparativePhysiologySynechococcus1995] and are expected for a
low light clade. The disjunct regions of the GAM plot results from
variable growth success of SS120 under 250 µM O~2~. Growth rate patterns
under red light were similar, although somewhat faster. In contrast,
under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak
PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol
photons m^-2^ s^-1^, indicated by the contour line labeled 0.41 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~ (Fig \@ref(fig:SS120PARGAM)B). Under red light and 25 µM O~2~ (Fig \@ref(fig:SS120PARGAM)B) SS120 grew across most conditions of peak PAR and
photoperiod, achieving fastest growth under long photoperiods and peak
PAR between 30 \~100 µmol photons m^-2^ s^-1^. Thus, the designation of
SS120 as a LL strain is dependent upon the [O~2~]. SS120 did not,
however, grow reliably under tested conditions at 2.5 µM O~2~.

@@

For MIT9313 under 25 µM O~2~, growth rate increased with increasing
photoperiods for all spectral wavebands tested (Fig \@ref(fig:MIT9313PARGAM)A), with the fastest overall growth rate for MIT9313
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "WW", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under peak PAR of 90 µmol photons m^-2^ s^-1^ and 16 h
white LED light photoperiod. In marked contrast to the 250 µM O~2~
growth experiments, MIT9313 grew when exposed to peak PAR of 180 µmol
photons m^-2^ s^-1^ and blue light under all photoperiods except 16 h;
additionally, white LED and red light treatments induced growth across
all tested photoperiods under 25 µM O~2~. The 25 µM O~2~, 4 h
photoperiod experiments under white LED light and were not performed due
to time constraints.

MIT9313 grew under 2.5 µM O~2~ particularly under blue LED light, albeit
generally slower than under the parallel experiments at 25 µM O~2~
(Fig \@ref(fig:MIT9313PARGAM)A). Growth estimates showed scatter among
replicates, suggesting 2.5 µM O~2~ is near the tolerance limit for
growth of MIT9313. Growth rates increased with longer photoperiods under
blue light treatments and peak PAR of 90 µmol photons m^-2^ s^-1^ but
did not grow under 16 h photoperiod. Growth for MIT9313 under PAR of 180
µmol photons m^-2^ s^-1^ and blue light treatment decreased with
increasing photoperiods with full growth inhibition under a 16 h
photoperiod. The red light peak PAR of 180 µmol photons m^-2^ s^-1^
showed similar growth rates to blue light for 8 and 12 h photoperiods.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 12, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^, achieved under 12 h blue light photoperiod and PAR of 90 µmol
photons m^-2^ s^-1^. The 2.5 µM O~2~ white LED treatments under 4, 8 and
16 h photoperiods and red light under 4 and 16 h photoperiods were not
performed as cultures were unlikely to grow.

@@

The GAM model in Fig \@ref(fig:MIT9313PARGAM)B summarizes MIT9313 growth
responses to red or blue peak PAR and photoperiod. Under
250 µM O~2~, Fig \@ref(fig:MIT9313PARGAM)B shows MIT9313 achieves
fastest growth rates between blue peak PAR of 30 µmol photons m^-2^
s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 12 h,
indicated by the contour line labeled 0.52 d^-1^ representing the 90^th^
percentile of achieved growth rates. Fig \@ref(fig:MIT9313PARGAM)B also
shows that growth rate increases with longer photoperiods, as long as
the blue peak PAR levels remain below 50 µmol photons m^-2^ s^-1^. In
contrast, under red light and 250 µM O~2~ MIT9313 grows faster while
exploiting higher peak PAR and longer photoperiods. Fig \@ref(fig:MIT9313PARGAM)B shows that MIT9313 can exploit all blue PAR levels
and most photoperiods with 90^th^ percentile of fastest growth rate
between 30 to 100 µmol photons m^-2^ s^-1^ PAR. Fig \@ref(fig:MIT9313PARGAM)B shows that MIT9313 maintains growth even under
2.5 µM O~2~, under photoperiods between 8 and 12 h and peak blue PAR
between 50 to 100 µmol photons m^-2^ s^-1^ PAR. Thus the designation of
MIT9313 as a LL clade is dependent upon [O~2~] and light spectra (Fig \@ref(fig:MIT9313PARGAM)B).

>>>>

Controlled growth experiments were performed using MCMIX-OD or MC1000-OD
PSI Multicultivators (Fig \@ref(fig:LabeledMC); PSI, Drásov, Czech
Republic). Each multicultivator individually controls 8 tubes at a
common temperature of 22°C. Each tube containing 70 mL of Pro99 media
was inoculated with 10 mL of growing maintenance culture approximately 16 to 22 hours before the start of each growth experiment. The tubes containing the cultures were placed in the Multicultivator water bath set at 22°C, sparged with the experimental [O~2~], and kept in the dark for approximately 12 hours until the photoregime of a sinusoidal photoperiod commenced the following morning such that peak PAR occurred at noon each day. In a factorial
matrix design, each tube was then subject to an individual combination
of sinusoidal photoperiod (4, 8, 12, 16 h); reaching a peak PAR (30, 90,
180 µmol photons m^-2^ s^-1^), with defined spectral bandwidth (White
LED, 660 nm, 450 nm). [O~2~] levels (2.5 µM, 25 µM, 250 µM) were imposed
by bubbling tubes with varying ratios of air and Nitrogen (N~2~), with
consistent 0.05% of Carbon Dioxide (CO~2~) gas, delivered through a 0.2
μm sterile microfilter via a G400 gas mixing system (Qubit Systems Inc.,
Kingston, Ontario, Canada). [O~2~] *in situ* was verified using oxygen
optodes (PyroScience, Germany) inserted into tubes for real-time
measurements, with a temperature probe in the bath of the bioreactor to
correct [O~2~] measures for temperature fluctuations. In addition, the
Pyroscience software corrected [O~2~] based on the salinity of the media
(32 ppt). The flow rate of the gas mixture was controlled, but
variations in bubbling speed, PAR and culture density affected the
[O~2~] achieved in each tube. A low [O~2~] of 0.5 µM - 5 µM (reported as
2.5 µM hereafter), was achieved by sparging with a gas mixture
containing 99.95% N~2~ and 0.05% CO~2~. An intermediate [O~2~] of 10 -
25 µM (reported hereafter as 25 µM) was achieved by sparging with a gas
mixture containing 98.95% N~2~, 0.05% CO~2~ and 1% O~2~. A high O~2~ of
200 µM - 280 µM (reported hereafter as 250 µM) was achieved by sparging
with lab air (78% N~2~, 21% O~2~, 1% Ar and 0.05% CO~2~).

@@

Guided in part by the evidence of ocean distributions of proteins from
*Prochlorococcus* we set up a matrix of photoperiods, PAR, spectral
bands, and [O~2~] to approximate current, and potential future,
latitudinal, depth and seasonal niches for *Prochlorococcus* strains. While we implemented measures to minimize the shock to cultures from exposure to experimental growth conditions by inoculating them the day before the experiment began and employing a sinusoidal photoperiod with a gradual increase in PAR exposure, we acknowledge that cells typically need multiple generations to acclimate to new growth conditions [@mooreCulturingMarineCyanobacterium2007]. As
mentioned, growth under red light could prove mechanistically
informative [@murphyPhotoinactivationPhotosystemII2017] to factors
limiting *Prochlorococcus* growth, we therefore included the red
spectral waveband even though it is not representative of
*Prochlorococcus* niches. Although *Prochlorococcus* is currently
limited to a narrow range of surface photoperiods, potential poleward
latitudinal expansions, in combination with attenuation of light with
depth, mean *Prochlorococcus* may potentially encounter a wide range of
photoperiods. Our growth rate determinations generally agree with those
from Moore *et al*. [@mooreComparativePhysiologySynechococcus1995], for
white LED and 250 µM O~2~, but our study is, to our knowledge, the first
to analyze the interactive growth responses of *Prochlorococcus* strains
to varying [O~2~], spectral wavebands and photoperiods.

@@

The GAM model in Fig \@ref(fig:SS120PARGAM)B summarizes growth responses
of SS120 to red or blue peak PAR and photoperiod, across the
2 imposed oxygen concentrations. Under 250 µM O~2~, Fig \@ref(fig:SS120PARGAM)B showed highest growth rates below blue light PAR of
50 µmol photons m^-2^ s^-1^ and photoperiods between 12 and 16 h,
indicated by the contour line labeled 0.35 d^-1^ (representing the
90^th^ percentile of achieved growth rate). Under 250 µM O~2~ SS120 is
constrained to deeper ocean waters through its intolerance of higher
blue PAR levels. These findings align with Moore *et al*.
[@mooreComparativePhysiologySynechococcus1995] and are expected for a
low light clade. The disjunct regions of the GAM plot results from
variable growth success of SS120 under 250 µM O~2~. Growth rate patterns
under red light were similar, although somewhat faster. In contrast,
under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak
PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol
photons m^-2^ s^-1^, indicated by the contour line labeled 0.42 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~ (Fig \@ref(fig:SS120PARGAM)B). Under red light and 25 µM O~2~ (Fig \@ref(fig:SS120PARGAM)B) SS120 grew across most conditions of peak PAR and
photoperiod, achieving fastest growth under long photoperiods and peak
PAR between 30 \~100 µmol photons m^-2^ s^-1^. Thus, the designation of
SS120 as a LL strain is dependent upon the [O~2~]. SS120 did not,
however, grow reliably under tested conditions at 2.5 µM O~2~.

@@

For MIT9313 under 25 µM O~2~, growth rate increased with increasing
photoperiods for all spectral wavebands tested (Fig \@ref(fig:MIT9313PARGAM)A), with the fastest overall growth rate for MIT9313
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "WW", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under peak PAR of 90 µmol photons m^-2^ s^-1^ and 16 h
white LED light photoperiod. In marked contrast to the 250 µM O~2~
growth experiments, MIT9313 grew when exposed to peak PAR of 180 µmol
photons m^-2^ s^-1^ and blue light under all photoperiods except 16 h;
additionally, white LED and red light treatments induced growth across
all tested photoperiods under 25 µM O~2~.

MIT9313 grew under 2.5 µM O~2~ particularly under blue LED light, albeit
generally slower than under the parallel experiments at 25 µM O~2~
(Fig \@ref(fig:MIT9313PARGAM)A). Growth rates showed scatter among
replicates, suggesting 2.5 µM O~2~ is near the tolerance limit for
growth of MIT9313. Growth rates increased with longer photoperiods under
blue light treatments and peak PAR of 90 µmol photons m^-2^ s^-1^ but
did not grow under 16 h photoperiod. Growth rates for MIT9313 under PAR of 180
µmol photons m^-2^ s^-1^ and blue light treatment decreased with
increasing photoperiods with full growth inhibition under a 16 h
photoperiod. The red light peak PAR of 180 µmol photons m^-2^ s^-1^
showed similar growth rates to blue light for 8 and 12 h photoperiods.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 12, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^, achieved under 12 h blue light photoperiod and PAR of 90 µmol
photons m^-2^ s^-1^. The 2.5 µM O~2~ white LED treatments under 4, 8 and
16 h photoperiods and red light under 4 and 16 h photoperiods were not
performed as cultures were unlikely to grow.

@@

The GAM model in Fig \@ref(fig:MIT9313PARGAM)B summarizes MIT9313 growth
responses to red or blue peak PAR and photoperiod. Under
250 µM O~2~, Fig \@ref(fig:MIT9313PARGAM)B shows MIT9313 achieves
fastest growth rates between blue peak PAR of 30 µmol photons m^-2^
s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 12 h,
indicated by the contour line labeled 0.52 d^-1^ representing the 90^th^
percentile of achieved growth rates. Fig \@ref(fig:MIT9313PARGAM)B also
shows that growth rate increases with longer photoperiods, as long as
the blue peak PAR levels remain below 50 µmol photons m^-2^ s^-1^. In
contrast, under red light and 250 µM O~2~ MIT9313 grows faster while
exploiting higher peak PAR and longer photoperiods. MIT9313 can exploit all blue PAR levels
and most photoperiods under 25 µM O~2~ with 90^th^ percentile of fastest growth rate
between 30 to 100 µmol photons m^-2^ s^-1^ PAR. MIT9313 maintains growth even under
2.5 µM O~2~, under photoperiods between 8 and 12 h and peak blue PAR
between 50 to 100 µmol photons m^-2^ s^-1^ PAR. Thus the designation of
MIT9313 as a LL clade is dependent upon [O~2~] and light spectra (Fig \@ref(fig:MIT9313PARGAM)B).
